---
title: "Bioinformatics Class 11"
author: "Natalie Toothacre"
date: "May 8, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## PDB Statistics

Download CSV file from PDB database <http://www.rcsb.org/stats/summary>. Read this into R and determine fration of the X-ray structures.

```{r}
pdb.stats <- read.csv("Data Export Summary.csv")
```

Lets calculate something

```{r}
percent <- (pdb.stats$Total / sum(pdb.stats$Total)) *100
names(percent) <- pdb.stats$Experimental.Method
percent
```


# Working with Bio3D

Lets load the package

```{r}
library(bio3d)
```

List the functions in this package and try an example

```{r}
#lbio3d()
#example(plot.bio3d)
```

Read a PDB file

```{r}
pdb <- read.pdb("1hsg")
pdb
```

Print subset of $atom data for the first two atoms

```{r}
pdb$atom[1:2, c("eleno", "elety", "x","y","z")]
```



Select and write out C alpha atom only PDB file

```{r}
inds <-  atom.select(pdb, "calpha")
inds
```

Check if we have what we want

```{r}
head(pdb$atom[inds$atom,])
```

Continue to trim our PDB and write it out

```{r}
ca.pdb <- trim.pdb(pdb, inds=inds)
ca.pdb

write.pdb(ca.pdb, file = "calpha_only.pdb")
```


```{r}
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)

pdbs <- pdbaln(files, fit = TRUE, exefile="C:/Users/Natalie/Downloads/muscle3.8.31_i86win32.exe")
```


##PCA of Adenylate kinase structures

find the sequence

```{r}
aa <- get.seq("1ake_A")
aa
```

Search the PDB database

```{r}
b <- blast.pdb(aa)
```


```{r}
hits <- plot(b)
```


```{r}

```










