---
title: "Writing a Function"
author: "Natalie Toothacre"
date: "May 7, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# New Function: plotbfactor 

Description:

The input to this funtion is the PDB ID of a protein of interest and requires the installation of the bio3d package.

This function pulls up the PDB file of the protein, selects for the atoms in chain A and with carbon alpha classification, and generates a plot of the b factor data from these atoms.

The output of this function is a plot of the b factor data.
  

```{r}

plotbfactor <- function(x) {
  
  #load the bio3d package
  library(bio3d)
  
  #load the PDB coordinate file of the protein of interest
  s <- read.pdb(x)
  
  #trim to chain A and Carbon alpha amino acid data
  s.chainA <- trim.pdb(s, chain = "A", elety = "CA")
  
  #call the atom data then the b factor data from s1.chainA
  s.b <- s.chainA$atom$b
  
  #plot the b factor data
  bf_plot <- plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor")
  
  return(bf_plot)
}

```


# Using the function

```{r}

plotbfactor("4AKE")

```



