---
title: "Bioinformatics Class 8"
author: "Natalie Toothacre"
date: "April 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)

```

```{r}
# Use the kmeans() function setting k to 2 and nstart=20
km <- kmeans(x, centers = 2, nstart = 20)
km

```

Q. How many points are in each cluster?
```{r}
km$size
```


Q. What ‘component’ of your result object details
 - cluster size?
 - cluster assignment/membership?
 - cluster center?
 
```{r}
#cluster assignment/membership
km$cluster

#cluster center
km$centers

```
 


 
```{r}
#Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

plot(x, col = km$cluster)
points(km$centers, col = "blue", pch = 15)

```
 
 
Q. Repeat for k=3, which one has the better total SS?

```{r}
km3 <- kmeans(x, centers = 3, nstart = 20)

km3$tot.withinss
```



## Hierarchical clustering

```{r}
# First we need to calculate point (dis)similarity as the Euclidean distance between observations
dist_matrix <- dist(x)

```

Lets look at the distance matrix

```{r}
class(dist_matrix)
```

```{r}
dim(as.matrix(dist_matrix))
```


```{r}
# The hclust() function returns a hierarchical clustering model
hc <- hclust(d = dist_matrix)

# the print method is not so useful here
hc 


```

Lets draw the tree

```{r}
plot(hc)

#draw a line at height h
abline(h=6, col="red")

#cuts by height h
cutree(hc, h=6)

#cuts into k groups
cutree(hc, k = 2)
```


Plotting data color coded by groups

```{r}
plot(x, col = cutree(hc, k = 4), pch = 16)
```



## Clustering Practice

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

```


```{r}
# Step 2. Plot the data without clustering
plot(x)

```


```{r}
# Step 3. Generate colors for known clusters (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)


```

Q. Use the dist(), hclust(), plot() and cutree() functions to return 2 and 3 clusters

```{r}
#create distance matrix
dist_matrix2 <- dist(x)

#get hierarchical clustering model
hc2 <- hclust(d = dist_matrix2)

#plot the model
plot(hc2)


```


```{r}
#cut tree at 2 groups
cutree(hc2, k = 2)

#cut tree at 3 groups
cutree(hc2, k = 3)
```


Plot graph with new col groups

```{r}
plot(x, col = cutree(hc2, k = 2))
```

```{r}
plot(x, col = cutree(hc2, k = 3))
```


Q. How does this compare to your known 'col' groups?

  The new groups overlap less


## PCA D:

Lets make some data

```{r}
## Initialize a blank 100 row by 10 column matrix
mydata <- matrix(nrow=100, ncol=10)
```

```{r}
## Lets label the rows gene1, gene2 etc. to gene100
rownames(mydata) <- paste("gene", 1:100, sep="") 
```

```{r}
## Lets label the first 5 columns wt1, wt2, wt3, wt4 and wt5
## and the last 5 ko1, ko2 etc. to ko5 (for "knock-out")
colnames(mydata) <- c( paste("wt", 1:5, sep=""),
 paste("ko", 1:5, sep="") )
```

```{r}
## Fill in some fake read counts
for(i in 1:nrow(mydata)) {
 wt.values <- rpois(5, lambda=sample(x=10:1000, size=1))
 ko.values <- rpois(5, lambda=sample(x=10:1000, size=1))

 mydata[i,] <- c(wt.values, ko.values)
}
head(mydata)
```

```{r}
#head(t(mydata))
```


```{r}
## lets do PCA
pca <- prcomp(t(mydata), scale=TRUE)

## See what is returned by the prcomp() function
attributes(pca) 
```

First two columns in pca$x correspond to PC1 and PC2
  Lets plot them

```{r}
## A basic PC1 vs PC2 2-D plot
plot(pca$x[,1], pca$x[,2]) 
```


The square of pca$sdev is variance

```{r}
## Variance captured per PC
pca.var <- pca$sdev^2 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pca.var.per
```

Make a scree plot from pca.var.per

```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")

#it is evident that most of the variation is accounted for by PC1
```

make plot more useful with this new data

```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)")) 

## Click to identify which sample is which
identify(pca$x[,1], pca$x[,2], labels=colnames(mydata))
```






