---
title: "Bioinformatics Class 7"
author: "Natalie Toothacre"
date: "April 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Funtions Again

Revisiting function from class 6

```{r}
source("http://tinyurl.com/rescale-R")
```

Lets see if we can use this function

```{r}
rescale(1:10)
```

Looks good lets break it

```{r}
#rescale(c(1:10, "string"))
```

Lets try new **rescale2()** function

```{r}
#rescale2(c(1:10, "string"))
```

## Write a NA cecking function

Here we write a new funtion to check NAs in two inputs

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

is.na(x)
```

```{r}
sum(is.na(x))
```

Lets put this together with `x` and `y`

```{r}
is.na(x) & is.na(y)
```

Our working snippet

```{r}
sum( is.na(x) & is.na(y))
```

This is the both_na() starting point

```{r}
function(x, y) {
  ## Check for NA elements in both input vectors 
  sum( is.na(x) & is.na(y) )
}
```

Test

```{r}
both_na(x,y)
```

And again

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x, y2)
```

Use stop() to give a more useful error code

```{r}
function(x, y) {
  ## Check for NA elements in both input vectors and don't allow re-cycling 
  if(length(x) != length(y)) {
    stop("Input x and y should be vectors of the same length", call.=FALSE)
  }
  sum( is.na(x) & is.na(y) )
}
```

Test

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na2(x, y2)
```

Refine and polish and test both_na3

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

both_na3(x, y)
```




## Another function example: gene intersection

```{r}
df1
df2

x <- df1$IDs
y <- df2$IDs
```

```{r}
x
y

```

```{r}
#intersect(x, y)
```

the **%in%** will be more helpful to us that **intersect()** 

```{r}
x %in% y

y %in% x
```

```{r}
x[x %in% y]
y[y %in% x]
```

we can now **cbind()** our results to yield an intersect

```{r}
cbind(x[x %in% y], y[y %in% x])
```

Make this into a funtion **gene_intersect**

```{r}
gene_intersect <- function(x, y) {
 cbind( x[ x %in% y ], y[ y %in% x ] )
}

gene_intersect(x, y)
```








