---
title: "Bioinformatics Class 13"
author: "Natalie Toothacre"
date: "May 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1000 Genomes Data Analysis

Downloaded genotype data from Ensemble for Asthma associated SNP rs8067378
See: (MXL) Mexican Ancestry in Los Angeles, California

```{r}
# Read CSV file
genotype <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378 (1).csv")
head(genotype)
```


Find how many of each genotype there are in this data set

```{r}
genotype.summary <- table(genotype[,2])
genotype.summary
```

Proportion of each genotype

```{r}
genotype.summary/nrow(genotype) * 100
```


> **Q5**: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL)
are homozygous for the asthma associated SNP (G|G)? 

Answer: 14%

## RNA-seq

Look mrore closely at sample HG00109 with GIG genotype for this SNP (From GBR population).

### Side note on FASTA quality scores 
```{r}
#install.packages("seqinr")
#install.packages("gtools")

library(seqinr)
library(gtools)

phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@")) - 33
phred
```

## RNA-seq population scale analysis

Read our results file from RNA-seq across ~230 samples

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
expr
```

Look at data
```{r}
table(expr$geno)
```

How many samples?
```{r}
nrow(expr)
```

Boxplot of expression levels of ORMDL3 by genotpe

```{r}
boxplot(exp ~ geno, data = expr, notch = TRUE)
```

Trying out ggplot

```{r}
#install.packages("ggplot2")
library(ggplot2)

ggplot(expr, aes(geno, exp)) + geom_boxplot()
```


```{r}
# Boxplot with the data shown
ggplot(expr, aes(geno, exp, fill=geno)) + 
  geom_boxplot(notch=TRUE, outlier.shape = NA) + 
  geom_jitter(shape=16, position=position_jitter(0.2), alpha=0.4)
```










